Case based Arabic Morphological Analysis with Compositional Non-deterministic Automata

### 2010-12-09 trunk/tex/rp.tex
evaluation

@@

We will build a query language for relational Arabic text mining 
and a front-end to compile a user query into an optimized
computation. 
We will check how user friendly and how expressive the language is.
We will compare our query language with the sophisticated 
interface of the MADA+TOKAN toolkit~\cite{Habash:09} and we will
look for other attempts in the industry .

\item Optimization methodologies
\begin{enumerate}
\item Use of language specific features 
\item Use of data in question specific features
\end{enumerate}
\item Entity and relational extractors
\item Statistical tools
\item Graph and text algorithms
\item Query solver
\item Case studies 
\begin{enumerate}
\item Hadith literature
\item Security
\item Health
\item Network content
\end{enumerate}
\end{enumerate}

\subsection{Evaluation Criteria}

The Lead PI will conduct biweekly meetings to track progress on the project.
 The team will maintain a dokuwiki based website to continuously post updates on the progress of the project.
 In Section "Research Design and Methods" we presented a schedule for delivering the different products of the project.
 The schedule can be used as a timeline for measuring the progress of the project.

The lexical and syntax analysis libraries will be compared with current open source libraries against commonly used corpora of data such as a collection of books or a collection of news reports.
 The dictionary module will be an augmentation over existing digitized dictionaries such as the Buckwalter~\cite{Tim04} morphological analyzer.
 In the entity and relational extraction tools we will use techniques native to the Arabic language and we will compare against English entity and relational extraction tools of similar or translated texts.
 We will also compare techniques that pre-process Arabic and then pass it to English IE and RE techniques.
 The logic solver will build over the IE and RE tools and will use off-the shelf quantifier and propositional solvers such as MiniSat~\cite{Een03}.
 The tagger is a GUI interface that allows fast manual inspection and refinement of the automated results.
 Each tool in the framework can start on its own with a prototypical implementation of the other needed tools.

Each tool in the framework stands alone and can operate on a well defined input and generate the desired output.
 The input of one tool, such as the entity extraction, can be automatically generated by another tool, such as the dictionary and the syntax analyzer.
  However, the input can also be manually generated if desired using the straightforward GUI tagger facility.
 Thus the tagger facility stands as a desired utility on its own and at the same time represents a contingency plan in case any module in the framework failed to deliver.

The team will publish the results of the research in the form of refereed conference and journal papers as well as tool papers addressing each of the tools.
 The team will also hold workshops to educate possible users on the tools once the tools are available.

@@

Once completed the Arabic text analysis framework will be an open source contribution and developers from the Arabic world in the open source community are expected to join efforts to keep it going.
 Furthermore, once completed the Arabic text analysis framework can sustain itself by charging on support or expert use.
 The framework or service on the framework can easily be commercialized to tend to the security, health, insurance, news, media, or literature business sectors.

@@

>>>>

The Lead PI will conduct biweekly meetings to track progress 
on the project.
The team will maintain a dokuwiki based website to continuously 
post updates on the progress of the project.

@@

We will build a query language for relational Arabic text mining 
and a front-end to compile a user query into an optimized
computation. 
We will check how user friendly and how expressive the language is.
We will compare our query language with the sophisticated 
interface of the MADA+TOKAN toolkit~\cite{Habash:09}.
We will look for other attempts to formalize Arabic text 
mining queries and we will also compare agains 
formal text mining queries from other text mining domains such
as the gene ontolgy domain~\cite{GeneOntology10},

\subsection{Optimizations}

We will develop and evaluate our two optimization methodologies
and look for other optimizations. 
First we will use language specific features based on the 
LCM of Arabic, the grammar of the query language, and 
the Arabic query provided by the user to simplify 
the LCM and the morphological analysis. 
Second we will look at language independent techniques
such as HMM and LSI models extracted from the 
text documents and use that to simplify the analysis. 
We will perform the analysis with and without the optimizations
and compare the results in terms of accuracy and efficiency.

%\subsection{ Entity and relational extractors}
\subsection{CATARM and case studies }

We will evaluate CATARM against at least three of the case studies
that we discussed previously and that are listed below. 
\begin{enumerate}
\item Hadith literature
\item Security
\item Health
\item Network content
\end{enumerate}

We anticipate CATARM to be able to successfully solve all the case
studies with resonable computational resources available to the 
average user. 
We will also generalize the components we will develop to solve
the case studies into a library of components that can be used
in other case studies.

\subsection{CATRAM evaluation }

The team will seek the assistance and help of an independent 
consultant in the middle of the second and third years of 
the duration of the project to evaluate and validate the results.

@@

Once completed the Arabic text analysis framework will be 
an open source contribution and developers from the Arabic 
world in the open source community are expected to join 
efforts to keep it going.
Furthermore, once completed the Arabic text analysis framework 
can sustain itself by charging on support or expert use.

The framework or service on the framework can easily be 
commercialized to tend to the security, health, insurance, market,
news, media, or literature business sectors.

@@

\subsection{Future work}

Enhance morphological analysis
Look for other optimizations
Generalize the framework to other Arabic-like languages
Generalize the case-based approach 
Integrate other text mining techniques such as vocalization,
suggestions...
